Compilation With Program Synthesis

### 2023-09-07 new-paper/semantics.tex
soundness
formatting

This subset of \tool does not have guards or \verb|if then else| expressions, but it does have pattern
matching. It also has the requirement that functions can only have one argument. Unlike the implementation, there
is no elaboration. As a result, every algebraic data type value must be lowered to a specific layout at every usage
and every function application must be explicitly instantiated with a layout for the argument and a layout for the result. We
also limit the available integer and Boolean operations for brevity.

@@

Typing rules for \tool expressions are given in
\autoref{fig:typing-rules}. These rules differ from standard typing
rules for a functional language due to the existence of layouts and
their associated constructors, like \verb|instantiate| and
\verb|lower|. If an expression is well-typed, then each use of
\verb|instantiate| and \verb|lower| only uses layouts together with
the ADT that they are defined for.

The rules also make use of a \textit{concreteness judgment}. The rules
for this judgment are given in \autoref{fig:concreteness-rules}. The
intuition of this judgment is that a type is ``concrete'' iff values
of that type can be directly represented in the heap machine
semantics. For example, an ADT type is \textit{not} concrete because a
layout has not been specified. However, once a particular layout is
specified for the ADT type, it becomes concrete. Base types, like
\verb|Int|, are also concrete.

@@

>>>>

This subset of \tool does not have guards or \lstinline[language=Pika]{if then else} expressions, but it does have pattern
matching. It also has the requirement that functions can only have one argument. Unlike the implementation, there
is no elaboration. As a result, every algebraic data type value must be lowered to a specific layout at every usage
and every function application must be explicitly instantiated with a layout for the argument and a layout for the result. We
also limit the available integer and Boolean operations for brevity.

@@

Typing rules for \tool expressions are given in
\autoref{fig:typing-rules}. These rules differ from standard typing
rules for a functional language due to the existence of layouts and
their associated constructors, like \lstinline[language=Pika]{instantiate} and
\lstinline[language=Pika]{lower}. If an expression is well-typed, then each use of
\lstinline[language=Pika]{instantiate} and \lstinline[language=Pika]{lower} only uses layouts together with
the ADT that they are defined for.

The rules also make use of a \textit{concreteness judgment}. The rules
for this judgment are given in \autoref{fig:concreteness-rules}. The
intuition of this judgment is that a type is ``concrete'' iff values
of that type can be directly represented in the heap machine
semantics. For example, an ADT type is \textit{not} concrete because a
layout has not been specified. However, once a particular layout is
specified for the ADT type, it becomes concrete. Base types, like
\lstinline[language=Pika]{Int}, are also concrete.

@@

\input{soundness.tex}
